Parsing Postfix log files

### 2009-02-16 doc/logparser-architecture.tex
Started actions section in architecture section, merging content from paper, and moving content to implementation chapter.

XXX TO BE WRITTEN\@.  START WITH THE CONTENT FROM THE PAPER, THEN CHECK THE
IMPLEMENTATION SECTION\@.

>>>>

XXX TO BE WRITTEN --- CHECK THE IMPLEMENTATION SECTION\@.

Each action is a separate procedure written to deal with a particular
category of input, e.g.\ rejections.  The actions are parser-specific: each
parser author will need to write the required actions from scratch unless
extending an existing parser.  It is anticipated that parsers based on this
architecture will have a high ratio of rules to actions, with the aim of
having simpler rules and clearer distinctions between the inputs parsed by
different rules.

The ability to add special purpose actions to deal with difficulties and
new requirements that are discovered during parser development is one of
the strengths of this architecture.  Instead of writing a single monolithic
function that must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.
Sometimes the new action will require the cooperation of other actions,
e.g.\ to set or check a flag.  There is a possibility of introducing
failure when modifying existing actions in this way, but the modifications
will be smaller and occur less frequently than with a monolithic
architecture, thus failures will be less likely to occur and will be easier
to test for and diagnose.  The architecture can be implemented in an object
oriented style, allowing sub-classes to extend or override actions in
addition to adding new actions; because each action is an independent
procedure, the sub-class need only modify the action it is overriding,
rather than reproducing large chunks of functionality.

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines there were many complications
to be overcome.  Some were the result of deficiencies in Postfix's logging
(some of which were rectified by later versions of Postfix); others were
due to the vagaries of process scheduling, client behaviour, and
administrative actions.  All were successfully accommodated in the Postfix
log parser: adding new actions was enough to overcome several of the
complications; others required modifications to a single existing action to
work around the difficulties; the remainder were resolved by adapting
existing actions to cooperate and exchange extra data (via the framework),
changing their behaviour as appropriate based on that extra data.

Actions may return a modified input line that will be parsed as if read
from the input stream, allowing for a simplified version of cascaded
parsing~\cite{cascaded-parsing}.  This powerful facility allows several
rules and actions to parse a single input, potentially simplifying both
rules and actions.

### 2009-02-16 doc/logparser-implementation.tex
Started actions section in architecture section, merging content from paper, and moving content to implementation chapter.

@@

All data extracted by the rule will be automatically saved, regardless of
the action which is invoked.

@@

The actions:

@@

THERE IS NEARLY A 1--1 MAPPING BETWEEN action AND postfix\_action.

>>>>

In the Postfix log parser developed for this project there are
\numberOFactions{} actions and \numberOFrules{} rules, with an uneven
distribution of rules to actions as shown in \graphref{Distribution of
rules per action}.  Unsurprisingly, the action with the most associated
rules is \texttt{DELIVERY\_REJECTED}, the action that handles Postfix
rejecting a mail delivery attempt; it is followed by \texttt{SAVE\_DATA},
the action responsible for handling informative log lines, supplementing
the data gathered from other log lines.  The third most common action is,
perhaps surprisingly, \texttt{UNINTERESTING}: this action does nothing when
executed, allowing uninteresting log lines to be parsed without causing any
effects (it does not imply that the input is ungrammatical or unparsed).
Generally rules specifying the \texttt{UNINTERESTING} action parse log
lines that are not associated with a specific mail, e.g.\ notices about
configuration files changing.  The remaining actions have only one or two
associated rules: some actions are required to address a deficiency in the
log files, or a complication that arises during parsing;  other actions
will only ever have one log line variant, e.g.\ all log lines showing that
a remote client has connected are matched by a single rule and handled by
the \texttt{CONNECT} action.

Using the \texttt{CONNECT} action as an example: it creates a new data
structure in memory for the new client connection, saving the data
extracted by the rule into it; this data will be entered into the database
when the mail delivery attempt is complete.  If a data structure already
exists for the new connection it is treated as a symptom of a bug, and the
action issues a warning containing the full contents of the existing data
structure, plus the log line that has just been parsed.

@@

Each action saves all data extracted by the rule's regex, if there is
enough information in the log line to identify the correct connection to
save the data to.

@@

@@

SEE TABLE (PROBABLY ON NEXT PAGE): THERE IS NEARLY A 1--1 MAPPING BETWEEN
action AND postfix\_action.  NEW ACTIONS COULD BE ADDED TO FORCE A 1--1
MAPPING, E.G. COMMIT + DISCARDED COULD BECOME A DISCARD ACTION\@.
MAIL\_QUEUED HAVING INFO AND PROCESSING CAN PROBABLY BE DEALT WITH, THE
BIGGEST DIFFICULTY WILL BE WITH SAVE\_DATA\@.
